---
title: "Assignment 7"
author: "Jack Scherer"
date: "2024-11-06"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
library(RColorBrewer)
library(latex2exp)
```

### Exercise 2 {-}

Using the `datasets::trees` data, complete the following. This question refreshes create a linear model, graphing the linear model, and introduces using some LaTeX expressions on the graph.

**a)** Create a regression model for $y=$ `Volume` as a function of $x=$ `Height`.
```{r}
model.trees <- lm( Volume ~ Height, data = trees )
```


**b)** Display the `summary` of the  model to view the y-intercept and slope of the regression line.
```{r}
summary(model.trees)
```


**c)** Using `ggplot2`, create a scatter plot of Volume vs Height. 
```{r}
plot.trees <- ggplot( data=trees, aes(x=Height, y=Volume) ) +
  geom_point()

plot.trees
```


**d)** Color the scatter using the `Girth` variable.
```{r}
plot.trees.2 <- ggplot( data=trees, aes(x=Height, y=Volume, color=Girth) ) +
  geom_point()

plot.trees.2
```


**e)** Modify the color scheme using a RColorBrewer palette.
```{r}
plot.trees.3 <- plot.trees.2 +
  #scale_color_distiller( palette = "RdYlBu" ) +
  scale_color_viridis_c(option='viridis')

plot.trees.3
```


**f)** Create a nice white filled rectangle to add text information. The following might be useful.

```{r, eval=FALSE}
annotate('rect', xmin=65, xmax=75, ymin=60, ymax=74,
         fill='white', color='black') +
```

```{r}
plot.trees.4 <- plot.trees.3 +
  annotate('rect', xmin=65, xmax=77, ymin=60, ymax=74,
         fill='white', color='black')

plot.trees.4
```


**g)** Use the `broom` package to extract the coefficients of the best-fit line. Add this information as an annotation to the graph, which should follow a form that looks like $\hat{y}_i = (INTERCEPT) + (SLOPE)* x_i$. Place the annotation within the white text box.
```{r}
broom::tidy( model.trees )

plot.trees.5 <- plot.trees.4 +
  geom_text( label="Volume = -87.12 + (1.54 * Height)", x=71, y=70, colour='black' )

plot.trees.5
```


**h)** Use the `broom` package to extract the coefficient of determination $r^2$ from the model. Add the annotation to your graph, which should look something like $R^2 = (VALUE)$

```{r}
broom::glance( model.trees )

plot.trees.6 <- plot.trees.5 +
  geom_text( label=latex2exp::TeX('$R^2 = 0.36$'), x=67.4, y=65, colour='black' )

plot.trees.6
```


**i)** Add the regression line in red. There are several ways to do this.
```{r}
trees <- trees %>% 
  mutate( fit = fitted(model.trees) )

plot.trees.7 <- plot.trees.6 +
  geom_line( aes( y=trees$fit ), colour='red' )

plot.trees.7
```


**j)** Properly label the axes of the graph.
```{r}
plot.trees.8 <- plot.trees.7 +
  labs( x='Tree Height', y='Tree Volume' )

plot.trees.8
```


**k)** Add a descriptive title to the graph.
```{r}
plot.trees.9 <- plot.trees.8 +
  labs( title='Tree Height vs. Volume' )

plot.trees.9
```




